SOM result tables

### 2025-01-14 analysis/Supplementary/supplementary_tables.qmd
Supplementary materials folder with supplementary tables. Removed supplementary tables from the main text, cleaned code

---
title: "Placeholder for article 3"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

# Background

## Aims

Our goal is to understand the provisioning, use and management strategies of chert lithic resources of the hunter-gatherer groups of Vale Boi during the Upper Paleolithic. We focus on the technological characteristics of the two broader groups of cherts identified at the site: local and non-local, based on the notion that local raw materials will show technological patterns different from those which are non-local. Expanding upon our previous works on chert types present throughout the UP sequence in the Terrace and Shelter area of Vale Boi, and their possible interpretations, and following the previously mentioned studies and models about raw material management and lithic technology organisation focusing mainly on the different management of local and non-local raw materials, associated with mobility and occupation length, we derive two main expectations (fig. 1):

1.  Local raw materials are expected to present a) low use-lives, with knapping strategies characterised by manufacture, use and discard; b) a decreased reduction intensity and low tools to debitage ratio; c) larger cores and lower blank-to-core ratio; d) less extensive reduction and reworking of tools; e) higher tool typology diversity. This may be related to the abundance of local raw materials which leads to less intensive life-use extension, but also to the residential characteristics of the site and low mobility, which increases the reliance on local raw materials.

2.  Non-local raw materials are expected to present a) a majority of formal tools, in the case of individual provisioning or higher use-lives, with knapping strategies characterised by the manufacture, use and maintenance; b) an increased reduction intensity and high tools to debitage ratio; c) smaller cores and higher blank-to-core ratio; d) more extensive reduction and high presence of retouch and/or tool maintenance; e) lower tool diversity, characterised by less specialised tools, such as the presence of retouched blanks as multipurpose tools. This may be related to scarcer or poorer quality of raw materials which leads to the prepared transportation of lithic resources to a site, but also related to more mobile settlement patterns as well as short-term site occupations.

Despite these expectations, it is important to notice that, especially in occupations with a high percentage of non-local cherts at the site (e.g., Gravettian occupations of level 7 and 6 of the Terrace at Vale Boi), these technological may not reflect a preoccupation with intensive management of a scarce and limited non-local resource, and be more a reflection of the type of provisioning applied following the suggested model by Kuhn (2004). In this sense, distance from source may have had little intrinsic meaning in the past, especially regarding non-local raw materials. There are several ways in which raw materials may be transported to a site, and the relationship between distance and cost may vary depending on how this takes place [@kuhn2004]. When non-local raw materials are as abundant as local ones, the difference in the technological organization between these materials may be less obvious (Andrefsky, 2004).

# Methods

# Results

```{r}
#| label: library

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(janitor)
library(ggpubr)
library(knitr)

@@

```{r}
#| label: terrace-databases-setup

cols_to_concat_terrace <- c("UNIT", "ID")
context$ID <- str_squish(context$ID) # Remove empty spaces from ID field
xyz$ID <- str_squish(xyz$ID) # Remove empty spaces from UNIT field

@@

```{r}
#| label: shelter-solutrean-setup

cols_to_concat_shelter <- c("YEAR", "ID") # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

@@

```{r}
#| label: general-tables

solutrean_general <- solutrean_dataset %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL", "INDET"))) %>% 
  tabyl(CLASS, RMTYPE) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

proto_general <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(CLASS != "Chip") %>% 
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TL", "INDET"))) %>% 
  tabyl(CLASS, RMTYPE) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

gravettian_general <- phase_dataset %>% 
  filter(Level %in% c(6, 7)) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TL", "INDET"))) %>% 
  tabyl(CLASS, RMTYPE) %>%
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(2) %>%
  adorn_ns("front")

@@

```{r}
#| label: tbl-general-gravettian
#| tbl-cap: Number and percentage of technotypological classes by raw material type from the Gravettian assemblage (levels 6 and 7) of the Terrace area of Vale Boi.

kable(gravettian_general)

```{r}
#| label: tbl-general-proto
#| tbl-cap: Number and percentage of technotypological classes by raw material type from the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area of Vale Boi.

kable(proto_general)

```{r}
#| label: tbl-general-solutrean
#| tbl-cap: Number and percentage of technotypological classes by raw material type from the Solutrean assemblage (levels A to C) of the Shelter area of Vale Boi.

kable(solutrean_general)

```{r}
#| label: raw-material-graph
#| include: false

library(ggplot2)
library(scales)

solutrean_dataset %>%
  filter(RMTYPE != "INDET") %>% 
  dplyr::group_by(LEVEL, RMTYPE) %>% 
  dplyr::summarise(n = n()) %>%
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(x = LEVEL, y = percent, fill = RMTYPE)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  #theme_minimal() + #type of theme
  labs(y = "Count %") +# change labels
  guides(fill=guide_legend(title=NULL)) + # remove legend title
  geom_text(aes(label=ifelse(percent >= 0.05, paste0(sprintf("%.1f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="white", size=3) +
  theme(axis.title.x = element_blank()) + 
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#8B6969", "#CDAF95", "#2B4F60",
                              "#996666", "#BDD2B6", "#798777", "#BDC7C9"))

## Assemblage description

For the Gravettian assemblage, local cherts show similar patterns, with T2 and T5 being the types with higher numbers of artefacts (n=110 and n=72, respectively). For all local cherts, blanks represent more than 50% of the artefacts. Cores and core maintenance products are only present in T2 and T5, although in small frequencies. For T2, cores represent 13.6% of the group and core preparation and maintenance products represent 1.8% (n=2). For T5 only 4 core fragments and 1 core preparation and maintenance product are present. Retouched tools seem to be especially relevant in T2 (\~18%), being the most frequent class after blanks. The presence of burin spalls (8.1%) in T2 also hints at the production of burins. In comparison, T5 shows similar percentages of retouched tools and shatter (\~12% and \~14%, respectively). T1 (chalcedony) shows similar patterns, although with higher percentages of shatter (34.3%). In comparison, non-local chert types (especially T6 with n=169 and T7 with n=86) show similar percentages of blanks (\~40-50%) but higher percentages of retouched tools (\~30%). Cores, core fragments and core preparation and maintenance products are also present even if in small percentages (\<4%), as well as shatter (\~12%). In both types burin spalls are also present (7.1% for T6 and 4.6% for T7). The TL cherts show, as expected, blanks and blank fragments (n=3 and n=2, respectively), and retouched tools (n=4). The presence of core maintenance products in TL is possibly related to the limited knapping of TL01 (jasper).

Regarding the Proto-Solutrean assemblage, the local types show high frequency of blanks (\>50%, including blank fragments), followed by shatter (\~11-27%), cores (\~9-12%) and retouched tools (\~5-15% ). A single burin spall was identified in T4. There is also the absence of core preparation and maintenance products across all local cherts. T1 (chalcedony) is composed mostly of blanks and a retouched tool (Vale Comprido point). The absence of cores but presence of shatter may hint at the knapping of this raw material at the site, but the absence of formal cores perhaps due to the quality of nodules/lenses, which are characterised by fractures. For non-local types, especially those with higher number of artefacts (T6-T8), blanks and blank fragments are the most frequent class (`r 38.3+10`%, `r 40.4+12.7`% and `r 46.1+23.1`%, respectively). For T6 the absence of cores but the presence of shatter and a single core preparation and maintenance product suggests knapping on site. Similarly, T7 shows the presence of cores and core fragments (`r 4.2+4.2`%), shatter (17.0%) and a single core preparation and maintenance product. In comparison, T8, despite more limited in number (n=13), shows only one core, with the rest of the artefacts being blanks (as previously noted) or retouched tools (23.1% and n=3). Aside from the TL artefacts, all non-local cores show at least 1 core and frequently the presence of shatter. As expected, the TL artifacts include a blank and a retouched. However, a shatter was also identified.

The Solutrean assemblages show similar patterns to those observed in previous assemblages. For local cherts, especially T2 (n=506) which accounts for the majority of local chert present in the studied samples, the majority of the debitage is composed of flakes and elongated blanks (`r 42+13.4+17.8+10.7`%). Cores are also present (10.5%), alongside a small percentage of core preparation and maintenance products (1.3%). In comparison to debitage products, retouched tools are scarce in the sample (\~3%). The non-local cherts most present in the Solutrean of the Shelter area are T6 (n=86) and T8 (n=30). These show similar patterns to the local cherts, with high percentages of blanks (\>50%). Cores and core preparation and maintenance are also present in small percentages (\<10%), as well as retouched tools (\<10%). However, these specific results are truncated by the lack of shatter, which were not present in the sample used for the study. It is important to notice, however, that alike previous results from the Solutrean assemblages of the Terrace area, a large portion of the studied sample was altered and did not allow for a chert type attribution (`r 201*100/914`%).

## Measurement and weight descriptive statistics

In the Gravettian assemblages, blanks and retouched tools show little differences in the measurements of artefacts of local and non-local chert types. The only exception are T6 and T7 blanks, which show slightly lower weight means and lower maximum length (\~35 mm) than the other types. For cores, T2 shows bigger (heavier) values than T6 (means of 21.3 and 12.6, and maximum weight of 37.2 gr and 22.5 gr, respectively), even if they seem to vary in morphology, with little patterns regarding thickness, width and length.

In the Proto-Solutrean assemblage, blanks show lower values for length and weight in non-local cherts (T6-T8), varying between a mean of \~19-20 mm and maximum of \<30 mm for length, and mean of \<3 gr and maximum of \<7 gr for weight. In retouched tools, weight also shows smaller means for non-local cherts (\~3-2 gr), although with higher standard deviation (\~1.6). Regarding cores, the sample size for non-local cherts is small and only T7 statistics could be obtained. Despite this, when compared to local cherts (T2-T5), T7 cores show smaller mean and maximum values (7.3 gr and 8 gr, respectively).

For the Solutrean assemblage, there seems to be no clear difference in the measurements and weights of artefacts between local and non-local cherts. This is true especially for the most represented classes (flakes and elongated blanks). Regarding cores, many of the local and non-local cherts are not available for the statistics due to missing data. However, the smallest weight mean seem to be for cores of T6 (16.5 gr), although with high standard deviation and maximum value (16.5 and 48.6 gr, respectively).

When comparing between technocomplexes, there are little differences in the general size and weight of the artifacts. For example, for key classes where size and weight differences could be expected (e.g., cores), the length measurements are mostly all located between 20-40 mm and weight between 10-40 gr. However, a few key differences were noted. T4 cores (the chert type most frequent in the Proto-Solutrean assemblage) shows the largest weights of cores of all assemblages, both in the Solutrean and Proto-Solutrean assemblage, as well as T5 in the Solutrean assemblage. Similarly, retouched tools show weights between 0.5-10 gr in all assemblages. T6 and T7 have consistently the smaller weights, especially compared to T2 during the Proto-Solutrean and Gravettian which shows slightly higher weights.

# Scatterplot for width and length
solutrean_scatter <- solutrean_stats |>
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) |>
  filter(MAXWIDTH !="0" & LENGTH != "0" & !RMTYPE %in% c("INDET","GB","TL")) |>
  dplyr::mutate(RMTYPE = case_when(
    RMTYPE == "TYPE 1" ~ "T1",
    RMTYPE == "TYPE 2" ~ "T2",
    RMTYPE == "TYPE 3" ~ "T3",
    RMTYPE == "TYPE 4" ~ "T4",
    RMTYPE == "TYPE 5" ~ "T5",
    RMTYPE == "TYPE 6" ~ "T6",
    RMTYPE == "TYPE 7" ~ "T7",
    RMTYPE == "TYPE 8" ~ "T8",
    RMTYPE == "TYPE 9" ~ "T9",
    RMTYPE == "TYPE 10" ~ "T10",
    RMTYPE == "TYPE 11" ~ "T11",
    TRUE ~ RMTYPE)) |>
   mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2", "T3", "T4", "T5", "T6","T7", 
                                        "T8", "T9", "T10", "T11"))) |>
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) |>
  ggplot(aes(x = LENGTH, y = MAXWIDTH, colour = RMTYPE)) +
  geom_point(alpha = 0.9, size = 3) +
  facet_wrap(~CLASS) +
  labs(x="Length (mm)", y="Max. Width (mm") +
  scale_colour_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6", "#BDC7C9"))

# Boxplot for individual measurements
# Length
solutrean_stats_length <- solutrean_stats %>%
  filter(CLASS %in% c("Flake", "ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(LENGTH !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, LENGTH, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y="Length (mm)") +
  #coord_cartesian(ylim =  c(0, 60)) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#8B6969", "#CDAF95", "#2B4F60",
                              "#996666", "#BDD2B6", "#798777", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), axis.title.x = element_blank())
# Weight
solutrean_stats_weight <- solutrean_stats %>%
  filter(CLASS %in% c("Flake", "ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types" ,y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
# Width
solutrean_stats_width <- solutrean_stats %>%
  filter(CLASS %in% c("Flake", "ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(MAXWIDTH !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, MAXWIDTH, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y="Width (mm)") +
  #coord_cartesian(ylim =  c(0, 60)) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#8B6969", "#CDAF95", "#2B4F60",
                              "#996666", "#BDD2B6", "#798777", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), axis.title.x = element_blank())

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_blanks <- solutrean_stats %>% 
  filter(CLASS %in% c("Flake", "ElongatedProd")) %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
solutrean_stats_cores <- solutrean_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
solutrean_stats_retouched <- solutrean_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

# Boxplots
# Width
proto_stats_width <- proto_stats %>%
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(MAXWIDTH !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, MAXWIDTH, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y="Width (mm)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#C74B50", "#D49B54", "#911F27", "#630A10",
                              "#FACE7F", "#8B6969", "#CDAF95", "#2B4F60",
                              "#996666", "#BDD2B6", "#798777", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), axis.title.x = element_blank())

# Scatterplots for width and length
proto_scatter <- proto_stats |>
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) |>
  filter(MAXWIDTH !="0" & LENGTH != "0" & !RMTYPE %in% c("INDET","GB","TL")) |>
  dplyr::mutate(RMTYPE = case_when(
    RMTYPE == "TYPE 1" ~ "T1",
    RMTYPE == "TYPE 2" ~ "T2",
    RMTYPE == "TYPE 3" ~ "T3",
    RMTYPE == "TYPE 4" ~ "T4",
    RMTYPE == "TYPE 5" ~ "T5",
    RMTYPE == "TYPE 6" ~ "T6",
    RMTYPE == "TYPE 7" ~ "T7",
    RMTYPE == "TYPE 8" ~ "T8",
    RMTYPE == "TYPE 9" ~ "T9",
    RMTYPE == "TYPE 10" ~ "T10",
    TRUE ~ RMTYPE)) |>
   mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2", "T3", "T4", "T5", "T6","T7", 
                                        "T8", "T9", "T10"))) |>
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) |>
  ggplot(aes(x = LENGTH, y = MAXWIDTH, colour = RMTYPE)) +
  geom_point(alpha = 0.9, size = 3) +
  facet_wrap(~CLASS) +
  labs(x="Length (mm)", y="Max. Width (mm") +
  scale_colour_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6"))

# Length
proto_stats_length <- proto_stats %>%
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(LENGTH !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, LENGTH, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(y="Length (mm)") +
  coord_cartesian(ylim =  c(0, 60)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), axis.title.x = element_blank())
# Weight
proto_stats_weight <- proto_stats %>%
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types", y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Scatterplots width and length
# Scatterplots for width and length
gravettian_scatter <- gravettian_stats |>
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) |>
  filter(MAXWIDTH !="0" & LENGTH != "0" & !RMTYPE %in% c("INDET","GB","TL")) |>
  dplyr::mutate(RMTYPE = case_when(
    RMTYPE == "TYPE 1" ~ "T1",
    RMTYPE == "TYPE 2" ~ "T2",
    RMTYPE == "TYPE 3" ~ "T3",
    RMTYPE == "TYPE 4" ~ "T4",
    RMTYPE == "TYPE 5" ~ "T5",
    RMTYPE == "TYPE 6" ~ "T6",
    RMTYPE == "TYPE 7" ~ "T7",
    RMTYPE == "TYPE 8" ~ "T8",
    RMTYPE == "TYPE 9" ~ "T9",
    RMTYPE == "TYPE 10" ~ "T10",
    TRUE ~ RMTYPE)) |>
   mutate(RMTYPE = factor(RMTYPE, levels = c("T1", "T2", "T3", "T4", "T5", "T6","T7", 
                                        "T8", "T9", "T10"))) |>
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) |>
  ggplot(aes(x = LENGTH, y = MAXWIDTH, colour = RMTYPE)) +
  geom_point(alpha = 0.9, size = 3) +
  facet_wrap(~CLASS) +
  labs(x="Length (mm)", y="Max. Width (mm") +
  scale_colour_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6"))

>>>>

---
title: "SOM result tables"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

@@

```{r}
#| label: databases-setup

# Terrace dataset setup
cols_to_concat_terrace <- c("UNIT", "ID")
context$ID <- str_squish(context$ID) # Remove empty spaces from ID field
xyz$ID <- str_squish(xyz$ID) # Remove empty spaces from UNIT field

@@

# Shelter dataset setup
cols_to_concat_shelter <- c("YEAR", "ID") # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

@@

# Blank-to-core ratios

# Solutrean ratio
solutrean_ratiobc <- solutrean_dataset %>% 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" | CLASS == "Thinning flake" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(RMTYPE != "INDET") %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Blank/Core" = Blank / Core)

# Proto-Solutrean ratio
proto_ratiobc <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>%
  mutate("Blank/Core" = Blank / Core)

# Gravettian ratio
gravettian_ratiobc <- phase_dataset %>% 
  filter(Level %in% c(6, 7)) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Blank/Core" = Blank / Core)

@@
